Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-10-11 Ben_Manuscripts/structure_paper/Final_Draft.tex
New TOC graphic
Final read-through edits

  More highly selective nanoporous membranes would be extremely useful 
  in performing complex aqueous separations with seawater and various 
  types of wastewater. For example, Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007} and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006} represent
  just a few of the diverse contaminants of water sources. By efficiently
  separating contaminants from feed solutions with highly selective membranes,
  it is possible to reduce the number of required membrane passes and post-treatment steps
  needed for a given filtration process \cite{werber_materials_2016}, thus
  lowering energy requirements. Additionally, one can also extract valuable resources
  from the feed streams. For example, flowback water produced during hydraulic
  fracturing of shale formations contains dissolved species such as acetate whose
  extraction has economic value \cite{dischinger_application_2017}.

@@

  Ordered, nanostructured porous membrane materials have become an increasingly popular
  research area for aqueous separations applications because they offer the 
  ability to control pore architecture at the molecular scale, thereby 
  permitting the design of solute-specific separation membranes 
  \cite{humplik_nanostructured_2011}. However, their development has been 
  limited by the ability to synthesize and scale various technologies. Graphene
  sheets are atomically thick, which results in excellent water permeability but
  defects during manufacturing severely impact selectivity \cite{cohen-tanugi_multilayer_2016}.
  Molecular dynamics (MD) simulations of carbon nanotubes have shown promise
  \cite{humplik_nanostructured_2011}, but synthetic techniques are unable to
  achieve scalable alignment and pore monodispersity
  \cite{hata_water-assisted_2004,maruyama_growth_2005}. Consequently, there is a
  need for new materials that could provide scalable nanostructured porous membranes.

@@

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.pdf}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.pdf}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{columns.pdf}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
	exhibits wedge-like character. (d) Monomers stack on top of each other to
	create columns with short range order then assemble into pores with  
	hydrophilic head groups (blue) facing towards the pore center. (e) The 
	pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

@@

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. Useful molecular-level modeling should incorporate a detailed picture of the
  nanoscopic pore structure, which is crucial to understanding the role of
  monomer structure in solute transport and membrane design. 
  Atomistic molecular dynamics simulations can provide the required level of detail
  (Figure \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can also observe how the choice of
  head group interacts with solutes of interest. In addition, we can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential.

  In this study, we achieve a  more realistic atomistic description of hexagonal LLC
  polymer membranes than, to our knowledge, has ever previously been created, and explore
  what new structural information can be gained and what structure hypotheses are
  supported by this model. We validate the results using as much experimental
  information as possible. We are most interested in reproducing the conclusions
  about structure drawn from small-angle X-ray scattering (SAXS) and wide-angle
  X-ray scattering (WAXS) experiments, as well as in matching ionic conductivity
  measurements \cite{feng_thin_2016}.

@@

  There are five major features of interest present in the 2D experimental
  pattern shown in Figure~\ref{fig:WAXS}.

@@

	\item \textit{R-$\pi$}: The location of the first is at $q_z$ = 1.7
	\AA$^{-1}$, corresponding to a real space separation of 3.7 {\AA}. Previous
	work attributes this reflection to $\pi$-$\pi$
	stacking between aromatic rings in the direction perpendicular to the membrane
	plane, or z-axis \cite{feng_scalable_2014}. For simplicity, we will refer to
	this reflection as R-$\pi$.

@@

	\item \textit{R-spots}: Within R-alkanes, are four spots of higher
	relative intensity.  Accordingly, we name these reflections R-spots. The
	location of all spots is $\sim37^{\circ}$ from the $q_z$ axis in their
	respective quadrants. In many liquid crystal systems, such spots are explained
	as the result of alkane chains tilted with respect to the membrane
	plane\cite{govind_simple_2001}.

@@

  	We are careful to avoid the term ``layers" since many liquid crystalline systems
 	have long range order in 1 or 2 spatial dimensions and short range order in 	
 	the other dimensions.~\cite{chaikin_principles_1995}  % page 58
 	In the system we are studying, there are long-range 2D correlations in the 
 	hexagonal array of pores (xy plane) and short range $z$-direction correlations within 
    stacked columns of monomers. In this study, we use atomistic molecular modeling
    to study how the system's structure is affected by the density of these monomer columns
    that surround each pore's hydrophilic core.

@@

	On the short timescales accessible to MD (even the 100's of nanoseconds of simulation 
	performed here are short compared to experimental timescales), we observe distinct metastable
	configurations which depend on starting configuration. We simulated XRD patterns of our system and 
	compared them to experimental 2D WAXS patterns (Figure~\ref{fig:WAXS}) so that we ensure our
	model creates a nanoscopic chemical environment maximally consistent with experiment within 
	the constraints	of our force field. Using this approach, we are able to confirm some previous
	interpretations	of the WAXS pattern and refute others.

@@

	While the Col\textsubscript{h} phase is described as dry, it is likely
	that small amounts of ambient water are absorbed into the system. The 
	hydrogen-bonding network formed by the water may play a role in structuring 
	the pore. We used simulated X-ray diffraction (XRD) patterns to uncover any 
	meaningful structural difference between a ``dry" and a ``wet" system.

@@

  Python scripts used to set up systems and conduct post-simulation trajectory analysis are
  available online at \texttt{https://github.com/bencoscia/llcsim}. The python scripts used
  to simulated XRD patterns are publicly available online at \\
  \texttt{https://github.com/joeyelk/MD-Structure-Factor}. Additional details are available in 
  Section \ref{S-section:python_scripts} of the SI.

@@

  We generated an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000 K to 50 K over 10
  ns. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration from the trajectory which we used for full system
  construction. Section \ref{S-section:parameterization} 
  of the Supporting Information (SI) provides further detail of the parameterization process.

@@

  The time scale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  requiring a more efficient way to build the system. Previous work has shown
  a united-atom model of LLC can self-assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (see Section \ref{S-section:self_assembly} of the 
  SI). To bypass the slow self-assembly process, we used a 
  variety of automated procedures to assemble monomers into a structure close to
  one of a number of hypothesized equilibrium configurations (See Figure 
  \ref{S-fig:build_procedure} of the SI). These initial structures may result
  in equilibration into slowly-intercoverting metastable states, an issue we will
  address later in the paper.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the equilibrium
  configuration is sensitive to some while insensitive to others. For example, we find the starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one chooses
  a reasonable value. The initial distance between pores, within a wide range, also has 
  little effect on the equilibrated structure. However, one should not start them
  too close or there will be high energy repulsions during early equilibration. We chose a
  pore radius of 0.5 nm and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the 
  experimental value of 4.12 nm for our initial configurations. A sensitivity analysis of 
  both parameters is presented in the SI, Section \ref{S-section:initial_config_dependence}. The 
  distance between vertically stacked monomers, the xy position of monomers with respect 
  to vertically adjacent monomers, and the number of columns per pore do influence the 
  equilibrated structure and require further justification for their choices. We rely on 
  experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. We rotated 
  each monomer so the plane of its aromatic head group would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with monomers placed so aromatic rings stack 3.7 \AA~apart in
  the $z$-direction. We also extensively explored a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this largest spacing yields non-physical behavior which is detailed in the 
  SI, Section \ref{S-section:initial_dbwl}.

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns per pore. See Figure 
  \ref{S-fig:stacking} of the SI for a detailed illustration 
  of the initial configurations in each mode.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. We located the pore centers by averaging the coordinates of sodium
  ions in their respective pores. We generated pore spacing statistics 
  using the bootstrapping technique (See Section \ref{S-section:p2p_stats} of the
  SI). 
  The pore spacings calculated in this way are consistent
  with one half of the $x$ and $y$ box vectors (See, Table \ref{S-table:p2p} in the
  SI). However, our method for their calculation does a better 
  job capturing the spread of pore-to-pore distances.

@@

  We normalized all diffraction patterns relative to R-alkanes. We believe that
  the alkane-alkane density, averaged over all angles, is the feature most likely
  to be replicated between experiment and simulation, as atomistic alkane
  force field parameters are relatively well-studied. Other features are dependent on system
  ordering which is likely to have some dependence on initial configuration.  We
  calculated the average intensity within R-alkanes of the experimental pattern,
  $I_{avg}$ and divided all intensities by this value. In this way, the average
  intensity of R-alkanes was set equal to 1. When calculating $I_{avg}$, we excluded
  intensities within $\pm$30\degree~of the meridional axis defined by $q_r=0$,
  since the simulated patterns differ from experiment in those regions in all
  cases (See Figure~\ref{fig:ralkanes}). Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$, as it appears in the simulated diffraction
  patterns, intersects with R-alkanes (See Figure~\ref{fig:XRDsim}). We set an
  upper bound on the colorbar by multiplying $I_{avg}$ by a scaling factor, $f$.
  Intensities that appear in the patterns $\geq$ $f\times I_{avg}$ are colored
  uniformly.  We applied the same scaling method to the simulated patterns. We
  chose a scaling factor of $f=3.1$ in order to visibly display all
  features in all patterns.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{ralkanes.pdf}  % Search WAXS.py for: fancy demo of r-alkanes normalization
  \caption{R-alkanes}\label{fig:ralkanes}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{rpi_rdouble.pdf}  %  Search WAXS.py for: R-pi / R-double intensity measurement illustration
  \caption{$q_z|_{r=0}$}\label{fig:rpi_rdouble}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{angular_integration.pdf} % Search WAXS.py for: Plot R-spots integrated intensity curve
  \caption{R-spots}\label{fig:rspots}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{saxs_xsection.pdf}  % 2DSAXS.py in Scripts folder 
  \caption{R-pores}\label{fig:saxs_xsection}
  \end{subfigure}
  \caption{(a) We measured the intensity of R-alkanes by calculating the average 
  intensity within the region bounded by $|\mathbf{q}|$ = 1.4~$\AA^{-1}$ and 
  1.57~$\AA^{-1}$ (between 4.0 and 4.5 \AA~in real space). We excluded intensities
  within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$, since the 
  simulated spectra overlap with R-$\pi$ in those regions in all cases. (b) We
  measured the intensity of R-$\pi$ and R-double based on the peaks of the $q_z$
  cross-section of the diffraction pattern. (c) We measured the intensity of R-spots
  by averaging the peak heights produced by radially integrating the patterns within 
  the R-alkanes region. We took the intensity of R-spots as the average of the 
  peak intensities near -37 and -143$\degree$. (d) We measured the intensity of 
  R-pores by measuring the height of the d\textsubscript{100} peak (the leading
  peak closest to $q_r$=0) of the cross-section of the diffraction patterns along
  the $q_r$ axis at $q_z$=0.} \label{fig:xrd_intensities}
  \end{figure}

@@

  We measured the one dimensional pair distribution function, $g(z)$, between centers 
  of masses (COM) of aromatic head group rings along the $z$-axis (perpendicular to
  the membrane plane). We averaged all 1D $z$-directional slices of the full 3D 
  correlation function within 2.1~\AA~of $(x, y)=(0, 0)$. We chose 2.1 \AA~as a crude 
  approximation of the radius of the phenyl ring plane. 
  We calculated the radius as the sum of the longest C-C distance within a phenyl 
  ring (2.8~\AA) and two times the carbon atom electronic radius (0.7~\AA)~\cite{slater_atomic_1964}.
  In this study, $g(z)$ is characterized by an oscillatory function with a period equal to the
  average distance between stacked monomers, and an amplitude that decays exponentially
  (see Figure~\ref{fig:correlation}). The rate of decay is related to the correlation 
  length, L, between monomer head groups. We estimated L by fitting the peaks of $g(z)$,
  using the python package \texttt{scipy.optimize.curve\_fit}~\cite{oliphant_python_2007},
  to a decaying exponential function of the form:
  \begin{equation}
  	Ae^{-z/L}
  	\label{eqn:decaying_exponential}
  \end{equation}
  where A is a fitting parameter for amplitude, $z$ is the independent variable of $g(z)$, 
  and L is the fit correlation length. We calculated the error in the estimated value
  of L as the square root of the diagonal entry of the covariance matrix of 
  optimized fit parameters.

@@

  We gave the simplified models the same amount of disorder present in the atomistic
  simulations. We observe two sources of disorder in our atomistic simulations: thermal motion of atoms
  during the simulation and quenched disorder created by rapid structural rearrangement during 
  early equilibration and is largely locked into place for the remainder of the 
  simulation, even at the 100 ns time scale. We measured thermal disorder by calculating the standard deviation of 
  the distribution of head group COM positions from their average positions.
  We measured quenched disorder by calculating the standard deviation of the distribution
  of head group COMs from their idealized average positions. In the 
  $z$-direction, we measured the deviation of the head group COMs from an
  equally spaced column of head groups. In the $xy$ plane, we calculated the standard 
  deviation in radial position from the pore center and the angular deviation from 
  equally spaced points surrounding the pore center. This method for calculating 
  quenched disorder inherently includes the convoluted thermal disorder.

@@

  R-alkanes appears close to its expected location. In all cases, the maximum
  intensity of R-alkanes along the $q_r$ axis at $q_z$=0, which we use to
  estimate its location, appears at a $|\mathbf{q}|$ value that is, at most,
  3.5\% higher than higher than experiment. The error in the simulated 
  diffraction patterns is less than the error due binning in each frequency
  direction, so was not further investigated.

@@

  \section*{Graphical TOC Entry}
  \begin{figure}
  \includegraphics[width=4.5cm]{toc_image_final.pdf}
  \includegraphics[width=4.07cm]{toc_waxs.pdf}
 \end{figure}

>>>>

  More highly selective nanoporous membranes would be extremely useful for
% BJC: in --> for
  performing complex aqueous separations with seawater and various types of
  wastewater. For example, Sodium chloride and boron in seawater
  \cite{fritzmann_state---art_2007} and organic micropollutants found in
  municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
  represent just a few of the diverse contaminants of water sources. By
  efficiently separating contaminants from feed solutions with highly selective
  membranes, it is possible to reduce the number of required membrane passes and
  post-treatment steps needed for a given filtration process
  \cite{werber_materials_2016}, thus lowering energy requirements. Additionally,
  one can also extract valuable resources from the feed streams. For example,
  flowback water produced during hydraulic fracturing of shale formations
  contains dissolved species such as acetate whose extraction has economic value
  \cite{dischinger_application_2017}.

@@

  Ordered, nanostructured porous membrane materials have become an increasingly popular
  research area for aqueous separation applications because they offer the
  %BJC: separations --> separation 
  ability to control pore architecture at the molecular scale, thereby
  permitting the design of solute-specific separation membranes
  \cite{humplik_nanostructured_2011}. However, their development has been limited
  by the ability to synthesize and scale various technologies. Graphene sheets
  are atomically thick, which results in excellent water permeability but defects
  during manufacturing severely impact selectivity
  \cite{cohen-tanugi_multilayer_2016}.  Molecular dynamics (MD) simulations of
  carbon nanotubes have shown promise \cite{humplik_nanostructured_2011}, but
  synthetic techniques are unable to achieve scalable alignment and pore
  monodispersity \cite{hata_water-assisted_2004,maruyama_growth_2005}.
  Consequently, there is a need for new materials that could provide scalable
  nanostructured porous membranes.

@@

  \begin{figure}
	\centering
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{NaGA3C11.pdf}
		\caption{}~\label{fig:monomer}
	\end{subfigure}
	\begin{subfigure}{.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{monomer_twocolor.pdf}
		\caption{}~\label{fig:atomistic_monomer}
	\end{subfigure}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
		\caption{}~\label{fig:wedge}
	\end{subfigure}
		\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{columns.pdf}
		\caption{}~\label{fig:wedge_layer}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
		\centering
		\includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
		\caption{}~\label{fig:hex_packing_simple}
	\end{subfigure}
	\caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
	exhibits wedge-like character. (d) Monomers stack on top of each other to
	create columns with short range order, then assemble into pores with 
        %BJC: order --> order, (added comma) 
	hydrophilic head groups (blue) facing towards the pore center. (e) The 
	pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

@@

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. Useful molecular-level modeling should incorporate a detailed
  picture of the nanoscopic pore structure, which is crucial to understanding the
  role of monomer structure in solute transport and membrane design.  Atomistic
  MD simulations can provide the required level of detail (Figure
  \ref{fig:detail}), assuming the force fields are sufficiently accurate.  With
  such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can also observe how the choice
  of head group interacts with solutes of interest. In addition, we can
  interchange counterions which may influence both the pore size and the strength
  of the Donnan potential.

  In this study, we achieve a more realistic atomistic description of a
  %BJC: of --> of a
  hexagonal LLC polymer membranes than, to our knowledge, has ever previously
  been created, and explore what new structural information can be gained and
  what structure hypotheses are supported by this model. We validate the results
  using as much experimental information as possible. We are most interested in
  reproducing the conclusions about structure drawn from small-angle X-ray
  scattering (SAXS) and wide-angle X-ray scattering (WAXS) experiments, as well
  as in matching ionic conductivity measurements \cite{feng_thin_2016}.

@@

  There are five major features of interest present in the 2D experimental
  WAXS pattern shown in Figure~\ref{fig:WAXS}.

@@

	\item \textit{R-$\pi$}: The location of the first is at $q_z$ = 1.7
	\AA$^{-1}$, corresponding to a real space separation of 3.7 {\AA}. Previous
	work attributes this reflection to $\pi$-$\pi$
	stacking between aromatic rings in the direction perpendicular to the membrane
	plane, or $z$-axis \cite{feng_scalable_2014}. For simplicity, we will refer to
	this reflection as R-$\pi$.

@@

	\item \textit{R-spots}: Within R-alkanes, are four spots of higher
	relative intensity.  Accordingly, we name these reflections R-spots. The
	location of all spots is $\sim37^{\circ}$ from the $q_r$ axis in their
	%BJC: q_z --> q_r
	respective quadrants. In many liquid crystal systems, such spots are explained
	as the result of alkane chains tilted with respect to the membrane
	plane\cite{govind_simple_2001}.

@@

	We are careful to avoid the term ``layers" since many liquid
	crystalline systems have long range order in 1 or 2 spatial dimensions and
	short range order in the other dimensions.~\cite{chaikin_principles_1995}  %
	page 58 In the system we are studying, there are long-range 2D correlations in
	the hexagonal array of pores ($xy$ plane) and short range $z$-direction
	correlations within stacked columns of monomers. In this study, we use
	atomistic molecular modeling to study how the system's structure is affected by
	the density of these monomer columns that surround each pore's hydrophilic
	core.

@@

	On the short timescales accessible to MD (even the 100's of nanoseconds
	of simulation performed here are short compared to experimental timescales), we
	observe distinct metastable configurations which depend on starting
	configuration. We simulated X-ray diffraction (XRD) patterns of our system and
	compared them to experimental 2D WAXS patterns (Figure~\ref{fig:WAXS}) so that
	we ensure our model creates a nanoscopic chemical environment maximally
	consistent with experiment within the constraints of our force field. Using
	this approach, we are able to confirm some previous interpretations	of the
	WAXS pattern and refute others.

@@

	While the Col\textsubscript{h} phase is described as dry, it is likely
	that small amounts of ambient water are absorbed into the system. The
	hydrogen-bonding network formed by the water may play a role in structuring the
	pore. We used simulated XRD patterns to uncover any meaningful structural
	difference between a ``dry" and a ``wet" system.

@@

  Python scripts used to set up systems and conduct post-simulation trajectory analysis are
  available online at \texttt{https://github.com/bencoscia/llcsim}. The python scripts used
  to simulated XRD patterns are publicly available online at \\
  \texttt{https://github.com/joeyelk/MD-Structure-Factor}. Additional details are available in 
  Section \ref{S-section:python_scripts} of the Supporting Information (SI).

@@

  We generated an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000 K to 50 K over 10
  ns. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration from the trajectory which we used for full system
  construction. Section \ref{S-section:parameterization} 
  of the SI provides further detail of the parameterization process.

@@

  The time scale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  requiring a more efficient way to build the system. Previous work has shown
  that a united-atom model of an LLC can self-assemble into the
  %BJC: show a united-atom --> that a united-atom ; of LLC --> of an LLC
  H\textsubscript{II} phase configuration in $\sim$1000 ns
  \cite{mondal_self-assembly_2013}. We attempted atomistic self-assembly by
  packing monomers into a box using Packmol \cite{martinez_packmol:_2009}.
  Simulations of greater than 100 ns show no indicators of progress towards an
  ordered system (see Section \ref{S-section:self_assembly} of the SI). To bypass
  the slow self-assembly process, we used automated procedures to
  %BJC: a variety of automated --> automated
  assemble monomers into a structure close to one of a number of hypothesized
  equilibrium configurations (See Figure \ref{S-fig:build_procedure} of the SI).
  These initial structures may result in equilibration into slowly-intercoverting
  metastable states, an issue we will address later in the paper.

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the
  equilibrium configuration is sensitive to some while insensitive to others. For
  example, we find the starting pore radius, defined as the distance of a chosen
  head group carbon from the pore's central axis, does not influence the
  equilibrium structure if one chooses a reasonable value. The initial distance
  between pores, within a wide range, also has little effect on the equilibrated
  structure. However, one should not start them too close or there will be high
  energy repulsions during early equilibration. We chose a pore radius of 0.5 nm
  and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the experimental
  value of 4.12 nm for our initial configurations. A sensitivity analysis of both
  parameters is presented in the SI, Section
  \ref{S-section:initial_config_dependence}. The distance between vertically
  stacked monomers, the $xy$ position of monomers with respect to vertically
  adjacent monomers, and the number of columns per pore do influence the
  equilibrated structure and require further justification for their choices. We
  rely on experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration
  based on R-$\pi$ and then allowed the system to readjust during equilibration.
  We rotated each monomer so the plane of its aromatic head group would be
  coplanar with the $xy$ plane. We explored three different initial monomer
  spacings. The first is exactly equal to R-$\pi$ with monomers placed so
  aromatic rings stack 3.7 \AA~apart in the $z$-direction. We also extensively
  explored a second system with an initial spacing of 5 \AA. We briefly explored
  a third system with an initial spacing of 10 \AA. However this largest spacing
  yields non-physical behavior which is detailed in the SI, Section
  \ref{S-section:initial_dbwl}.

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns per pore. See Figure 
  \ref{S-fig:stacking} of the SI for a detailed illustration 
  of the initial configurations in each mode.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. We located the pore centers by averaging the coordinates of
  sodium ions in their respective pores. We generated pore spacing statistics
  using the bootstrapping technique (See Section \ref{S-section:p2p_stats} of the
  SI). The pore spacings calculated in this way are consistent with one half of
  the $x$ and $y$ box vectors (See Table \ref{S-table:p2p} in the SI). However,
  our method for their calculation does a better job capturing the spread of
  pore-to-pore distances.

@@

  We normalized all diffraction patterns relative to R-alkanes. We believe that
  the alkane-alkane density, averaged over all angles, is the feature most likely
  to be replicated between experiment and simulation, as atomistic alkane force
  field parameters are relatively well-studied~\cite{wang_development_2004}.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration.  We calculated the average intensity
  within R-alkanes of the experimental pattern, $I_{avg}$, and divided all
  intensities by this value. In this way, the average intensity of R-alkanes was
  set equal to 1. When calculating $I_{avg}$, we excluded intensities within
  $\pm$30\degree~of the meridional axis defined by $q_r=0$, since the simulated
  patterns differ from experiment in those regions in all cases (See
  Figure~\ref{fig:ralkanes}). Specifically, in contrast to the experimental WAXS
  pattern, R-$\pi$, as it appears in the simulated diffraction patterns,
  intersects with R-alkanes (See Figure~\ref{fig:XRDsim}). We set an upper bound
  on the colorbar by multiplying $I_{avg}$ by a scaling factor, $f$.  Intensities
  that appear in the patterns $\geq$ $f\times I_{avg}$ are colored uniformly.  We
  applied the same scaling method to the simulated patterns. We chose a scaling
  factor of $f=3.1$ in order to visibly display all features in all patterns.

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{ralkanes.pdf}  % Search WAXS.py for: fancy demo of r-alkanes normalization
  \caption{R-alkanes}\label{fig:ralkanes}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{rpi_rdouble.pdf}  %  Search WAXS.py for: R-pi / R-double intensity measurement illustration
  \caption{$q_z|_{r=0}$}\label{fig:rpi_rdouble}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{angular_integration.pdf} % Search WAXS.py for: Plot R-spots integrated intensity curve
  \caption{R-spots}\label{fig:rspots}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
  \centering
  \includegraphics[width=\textwidth]{saxs_xsection.pdf}  % 2DSAXS.py in Scripts folder 
  \caption{R-pores}\label{fig:saxs_xsection}
  \end{subfigure}
  \caption{(a) We measured the intensity of R-alkanes by calculating the average 
  intensity within the region bounded by $|\mathbf{q}|$ = 1.4~$\AA^{-1}$ and 
  1.57~$\AA^{-1}$ (between 4.0 and 4.5 \AA~in real space). We excluded intensities
  within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$, since the 
  simulated spectra overlap with R-$\pi$ in those regions in all cases. (b) We
  measured the intensity of R-$\pi$ and R-double based on the peaks of the $q_z$
  cross-section of the diffraction pattern. (c) We measured the intensity of R-spots
  by averaging the peak heights produced by radially integrating the patterns within 
  the R-alkanes region. We took the intensity of R-spots as the average of the 
  peak intensities near 37 and 143$\degree$. (d) We measured the intensity of 
  R-pores by measuring the height of the d\textsubscript{100} peak (the leading
  peak closest to $q_r$=0) of the cross-section of the diffraction patterns along
  the $q_r$ axis at $q_z$=0.} \label{fig:xrd_intensities}
  \end{figure}

@@

  We measured the one dimensional pair distribution function, $g(z)$, between centers 
  of masses (COMs) of aromatic head group rings along the $z$-axis (perpendicular to
  the membrane plane). We averaged all 1D $z$-directional slices of the full 3D 
  correlation function within 2.1~\AA~of $(x, y)=(0, 0)$. We chose 2.1 \AA~as a crude 
  approximation of the radius of the phenyl ring plane. 
  We calculated the radius as the sum of the longest C-C distance within a phenyl 
  ring (2.8~\AA) and two times the carbon atom electronic radius (0.7~\AA)~\cite{slater_atomic_1964}.
  In this study, $g(z)$ is characterized by an oscillatory function with a period equal to the
  average distance between stacked monomers, and an amplitude that decays exponentially
  (see Figure~\ref{fig:correlation}). The rate of decay is related to the correlation 
  length, L, between monomer head groups. We estimated L by fitting the peaks of $g(z)$,
  using the python package \texttt{scipy.optimize.curve\_fit}~\cite{oliphant_python_2007},
  to a decaying exponential function of the form:
  \begin{equation}
  	Ae^{-z/L}
  	\label{eqn:decaying_exponential}
  \end{equation}
  where A is a fitting parameter for amplitude, $z$ is the independent variable of $g(z)$, 
  and L is the fit correlation length. We calculated the error in the estimated value
  of L as the square root of the diagonal entry of the covariance matrix of 
  optimized fit parameters.

@@

  We gave the simplified models the same amount of disorder present in the
  atomistic simulations. We observe two sources of disorder in our atomistic
  simulations: thermal motion of atoms during the simulation and quenched
  disorder created by rapid structural rearrangement during early equilibration
  that is largely locked into place for the remainder of the simulation, even at
  %BJC: and --> that
  the 100 ns time scale. We measured thermal disorder by calculating the standard
  deviation of the distribution of head group COM positions from their average
  positions. We measured quenched disorder by calculating the standard deviation
  of the distribution of head group COMs from their idealized average positions.
  In the $z$-direction, we measured the deviation of the head group COMs from an
  equally spaced column of head groups. In the $xy$ plane, we calculated the
  standard deviation in radial position from the pore center and the angular
  deviation from equally spaced points surrounding the pore center. This method
  for calculating quenched disorder inherently includes the convoluted thermal
  disorder.

@@

  R-alkanes appears close to its expected location. In all cases, the maximum
  intensity of R-alkanes along the $q_r$ axis at $q_z$=0, which we use to
  estimate its location, appears at a $|\mathbf{q}|$ value that is, at most,
  3.5\% higher than experiment. The error in the simulated diffraction patterns
  is less than the error due binning in each frequency direction, so was not
  further investigated.

@@

  \section*{Graphical TOC Entry}
  \begin{figure}
  \includegraphics[width=3.5in]{new_toc_graphic.pdf}
%  \includegraphics[width=4.5cm]{toc_image_final.pdf}
%  \includegraphics[width=4.07cm]{toc_waxs.pdf}
 \end{figure}
